---
title: "week4-E4_demo"
author: "AureliaMo"
date: "2023-05-11"
output: html_document
---

# Demo

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)

library(sf)
library(tmap)

# library(SimilarityMeasures)
```

how to write a function in r
```{r}

testfun <- function(){
  print("this function doesn't do anything")
}
testfun()

testfun <- function(sometext){
  print(sometext)
}
testfun("this function doesn't do anything") # doesnt work

class(testfun)

```


```{r}

my_age <- function(birthday, unit="days"){
  now <- Sys.time()
  difftime(now, birthday, units = unit)
}

my_age("1993-07-08", "mins")

# careful with the variable names you use in the function. always specify the variables within the function
```

## Task 1: Write your own functions

create a function for euclidean distance

```{r}


euclid <- function(x, y){
  sqrt((lead(x)-x)^2 + (lead(y)-y)^2)
}

wildschwein <- read_delim("wildschwein_BE_2056.csv")

wildschwein |> 
  mutate(steplength = euclid(E, N))

```
`

```{r}
euclid <- function(x, y){
  steplength <- sqrt((lead(x)-x)^2 + (lead(y)-y)^2)
}

euclid(x= c(0,1), y=c(2,4))
# doesn't return anything, because the result is stored somewhere


euclid <- function(x, y){
  steplength <- sqrt((lead(x)-x)^2 + (lead(y)-y)^2)
  return(steplength)
}

euclid(x= c(0,1), y=c(0,1))
```

## Task 2: prepare Analysis

In the next tasks we will look for “meet” patterns in our wildboar data. To simplify this, we will only use a subset of our wildboar data: The individuals Rosa and Sabi for the timespan 01.04.2015 - 15.04.2015. 

Meet: at the same time at the same place

```{r}
wildboar2<- wildschwein |> 
  filter(TierName %in% c("Rosa", "Sabi")) |> 
  # filter(TierName == "Rosa" | TierName == "Sabi")# its the same
  filter(DatetimeUTC >= as.POSIXct("2015-04-01 00:00:00", tz = "UTC")) |> 
  filter(DatetimeUTC <= as.POSIXct("2015-04-14 23:59:59", tz = "UTC"))

wildboar2 |> 
  ggplot(aes(E,N, colour=TierName))+
  geom_point( alpha = .2)+
  coord_equal()
```

## Task 3: Create Join Key
now wee need to know when two animals were sampled at the same time
we can do this by joining. but first we need to normalize the times

```{r}
wildboar2 <- wildboar2 |> 
  select(-TierID, -CollarID) |> #remove unnecessary columns
  mutate(Datetime_round = lubridate::round_date(DatetimeUTC, "15 minutes")) # rounds the times to the nearest 15 minutes

rosa <- wildboar2 |> 
  filter(TierName=="Rosa")

sabi <- wildboar2 |> 
  filter(TierName=="Sabi")

# the sampling time is a little different. rosa was sampled at 00:00:10 and sabi at  00:00:11. so to use time as a join key, we need to somehow round or roll the times

rosa_sabi <- inner_join(sabi, rosa, by="Datetime_round", suffix = c("_sabi", "_rosa"))

meets <- rosa_sabi |> 
  mutate(
    distance = sqrt((E_sabi-E_rosa)^2 + (N_sabi-N_rosa)^2),
    meet = distance < 100
  )

# to calculate the distance here we would need another function without lead()
# 
# euclid2 <- function(x, y, x2, y2){
#   steplength <- sqrt((x-x2)^2 + (y-y2)^2)
# }
# 
# rosa_sabi |> 
#   mutate(
#     distance = euclid2(E_sabi, N_sabi, E_rosa, N_rosa),
#     meet = distance < 100
#   )
```


## Task 5: Visualize data


```{r}

# filter for meets

meets <- meets |> 
  filter(meet==TRUE)

ggplot(rosa, aes(E,N, colour=TierName), alpha = .2)+
  geom_point(sabi, aes(E,N, colour=TierName), alpha = .2)+
  geom_point(meets, aes(E_sabi, N_sabi))+
  coord_equal()


```

